### 2020-09-16 ieee_temp_groll/ieee_groll.tex

\title{Ground-roll noise attenuation using a simple and effective approach based on local bandlimited orthogonalization}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\author{\IEEEauthorblockN{Yangkang Chen$^1$, Shebao Jiao$^2$, Jianwei Ma$^3$, Hanming Chen$^2$, Yatong Zhou$^4$, and Shuwei Gan$^2$}\\
$^1$\IEEEauthorblockA{Jackson School of Geosciences,
The University of Texas at Austin,
University Station, Box X,
Austin, TX 78713-8924, USA,
Email: ykchen@utexas.edu}\\
$^2$\IEEEauthorblockA{State Key Laboratory of Petroleum Resources and Prospecting, 
China University of Petroleum, 
Fuxue Road 18th,
Beijing, China, 102200, 
Email: gsw19900128@126.com \& huichanming@126.com} \\
$^3$ \IEEEauthorblockA{Department of Mathematics,
Harbin Institute of Technology,
Harbin, China, 
jma@hit.edu.cn}\\
$^4$\IEEEauthorblockA{School of Electronic and Information Engineering,
Hebei University of Technology,
Xiping Road No. 5340, Beichen District,
Tianjin, China, 300401,
zyt@hebut.edu.cn}\\
}
\maketitle
%512375919@qq.com

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	Defined by myself 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
\newcommand{\dlo}[1]{}
\newcommand{\wen}[1]{#1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	Defined by myself 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
 The bandpass filtering is a common way for estimating the ground-roll noise in land seismic data because of the low-frequency property of ground-roll noise. However, the frequency-mixing problem makes bandpass filtering inconvenient or even impossible to achieve a successful removal of ground-roll noise.  In this paper, we propose a novel band-limited orthogonalization approach for removing ground-roll noise without harming useful primary reflections. We orthogonalize the initially denoised signal via bandpass filtering with a relatively high low-bound-frequency (LBF), and the corresponding noise section that contains loss of primary reflections, using local signal-and-noise orthogonalization.  The local orthogonalization guarantees that the least coherent primary reflections are lost in the noise section. The procedure of the proposed approach is very convenient to implement because only a bandpass filtering and a regularized division between the initially denoised signal and initial noise are used. An open-source field dataset demonstrates a successful performance using the proposed approach. 
\end{abstract}

\begin{keywords}
Ground-roll noise, bandpass filtering, local orthogonalization, regularized division
\end{keywords}

\section{Introduction} 
Ground-roll noise is a type of seismic noise, with a high amplitude, low frequency, and low velocity. Ground-roll noise is the main type of coherent noise for land seismic \dlo{survey}\wen{surveys}. Ocean bottom node (OBN) based marine seismic surveys may also contain this type of noise \cite{yangkang20142}. The ground-roll noise is a type of source-generated noise and can be scattered by near-surface heterogeneities.  The ground-roll noise usually mask the shallow reflections at short offset, and deep reflections at larger offset \cite{claerbout1983,saatilar1988,henley2003}, and must be removed before the subsequent processing tasks. Instead of being incoherent along the spatial direction like random noise \cite{wencheng2015asa,yangkang2015}, the ground-roll noise is \wen{usually} coherent \wen{along the spatial dimension} \dlo{and behave much similarly to the primary reflections}, which makes the removal of them very difficult using simple signal processing methods. 
There have been a lot of researches about removing ground-roll noise done in the literature and many researchers have proposed different methods for handling the ground-roll noise problem \cite{shieh1990,brown2000}. Most of the ground-roll noise removal approaches either fail to remove all the ground-roll noise or remove much useful primary reflections energy. An efficient and effective technique for removing ground-roll noise is always in demand.

The simplest and most straightforward way for removing ground-roll noise might be bandpass filtering. Because of the low-frequency property, a simple high-pass filter is usually applied to the seismic record to attenuate the ground-roll noise. However, the low-bound-frequency (LBF) for the bandpass filter is not easy to choose, because higher LBF will cause much damage to primary reflections while lower LBF will fail to remove enough ground-roll noise. This issue exists because of the coupling of primary reflections and ground-roll noise in the frequency bands. One way to solve this problem is to use matching filtering. The low-pass data is used as an initial guess for the ground-roll noise and then a least squares (LS) based matching filter is calculated to match the initial ground-roll noise to the raw seismic data best \cite{carson2006,stephen2007,halliday2010}. The matched ground-roll noise is then subtracted from raw seismic data to obtain the output. This adaptive subtraction method highly depends on the initial prediction of the ground-roll noise. Besides, in the case of highly non-stationary primary reflections and ground-roll noise, a conventional stationary matching filtering will fail to obtain an acceptable result.

%\cite{james1968} proposed the multichannel stacking filters for attenuating those types of noise which \dlo{occur}\wen{occurred} at relatively well-determined points on neighbor traces, such as the ground-roll noise. \cite{prasad1980} studied and compared the performance of the least-mean-squares (LMS) adaptive filter, the adaptive lattice filter, and the adaptive Kalman filter identifier.  \cite{cahit1983} proposed an effective surface wave removal approach termed as vibroseis whitening (VSW) approach, which was tailed specifically for vibroseis data. The VSW approach was based on the application of time-varying amplitude scaling before vibroseis crosscorrelation. This scaling resulted in a signal-to-noise ratio (SNR) improvement equal\dlo{s} to the gain expected from crosscorrelation. This approach \dlo{is}\wen{was} also termed as the spectral balancing approach. \cite{ruhi1988} proposed a one-dimensional, linear frequency-modulated matched filter for removing ground-roll noise. This approach works well when the frequency bands of seismic reflections and ground-roll noise overlap. \cite{carson2006} proposed to use curvelet transform to remove ground-roll noise by a two-step process. The first step is to identify the ground-roll noise from the seismic data using an appropriate approach and then separate the signal and ground-roll noise using a block-coordinate relaxation method that seeks the sparest set for weighted curvelet coefficients of the ground-roll noise and the sought-after reflectivity. \cite{ruhi1994} proposed an adaptive lattice filter for attenuating ground-roll noise. The lattice filter has the same order of computational complexity as the LMS adaptive filter but enjoys a faster convergence rate.  \wen{\cite{regone1998} and \cite{ozbek2000} proposed to use acquisition-based suppression schemes to attenuate the ground-roll noise based on the use of recording arrays. \cite{stephen2007} proposed a polarization filtering approach to attenuate ground-roll noise, in which they combined statics, signal-to-noise ratio, and eigenimages to construct the ground-roll noise model, followed by adaptive subtraction.}  \wen{There is a type of ground-roll noise attenuation approaches that are based on the seismic interferometry \cite{halliday20072,halliday20092}.} \cite{yanwei2009} proposed an interferometric method to predict and subtract surface waves in seismic data. Instead of the traditional frequency-based one-step prediction, the interferometric method uses a multi-prediction and subtraction strategy to attenuate more ground-roll noise and preserve more useful energy. However, the implementation requires a careful parameters designing and tuning process. \wen{\cite{halliday2010} also proposed an interferometric ground-roll removal approach. They used seismic interferometry to estimate scattered ground-roll noise between pairs of receiver locations.}   A recently proposed approach by \cite{liuyang2012} utilized a time-frequency decomposition technique to separate ground-roll noise and primary reflections. The basic principle of this method is that instead of using a global frequency difference between the primary reflections and ground-roll noise, it uses the local time-frequency property of ground-roll noise to achieve the noise free data. 
In this paper, we proposed a simple but effective way for removing all the ground-roll noise without harming useful reflections. We first apply a simple bandpass filtering to raw seismic data in order to remove all the ground-roll noise (thus the LBF should be a little bit high to guarantee no ground-roll noise left in the data), then because of loss of useful primary reflections, we use local signal-and-noise orthogonalization proposed recently by \cite{yangkang2015ortho} to orthogonalize the primary reflections and ground-roll noise in order to retrieve the lost useful signals back to the initial bandpass filtered data. The performance of the proposed approach on the pre-stack dataset shows successful performance.

\section{Method}
\subsection{Bandpass filtering and its frequency-mixing problem}
A bandpass filtering process can be simply summarized as:
\begin{equation}
\label{eq:bandpass}
\mathbf{\hat{d}} = \mathbf{F}^{-1} \mathbf{B}_{fl,fh} \left(\mathbf{F} \mathbf{d}\right),
\end{equation}
where $\mathbf{d}$ and $\mathbf{\hat{d}}$ denote the unfiltered and filtered data, respectively. $\mathbf{F}$  and $\mathbf{F}^{-1}$ denotes a pair of forward and inverse Fourier transforms. $\mathbf{B}_{fl,fh}$ denotes the bandpass filter with a high-bound-frequency (HBF) $fh$, and low-bound-frequency (LBF) $fl$. For the purpose of removing ground-roll noise, $fh$ is usually chosen as Nyquist frequency, the only parameter to choose is only the LBF.

The problem of bandpass filtering for removing ground-roll noise is the inconvenience, or impossibility to choose an optimal LBF, because of the frequency-mixing problem of ground-roll noise and primary reflections. Figure \ref{fig:field-1-0,dif-1-0,field-2-0,dif-2-0} shows a demonstration of the frequency-mixing problem. When $fl=25$ Hz, all the ground-roll noise have been removed, and denoised section (Figure \ref{fig:field-1-0}) does not contain any ground-roll noise. However, the noise section (Figure \ref{fig:dif-1-0}) contains a lot of coherent signals: both direct waves and primary reflections. When $fl=10$ Hz, the noise section (Figure \ref{fig:dif-2-0}) does not contains any coherent reflection or direct waves, however, the denoised section (Figure \ref{fig:field-2-0}) still contains a large amount of ground-roll noise.

One of the most commonly used approaches to solve the frequency-mixing problem is to use matching filtering. The removed ground-roll noise after a common bandpass filtering is used as an initial guess for the ground-roll noise. A least squares (LS) based matching filter is then calculated to match the initial ground-roll noise to the raw seismic data based on the least-energy assumption. The matched ground-roll noise is then subtracted from the raw seismic data to obtain the ground-roll noise attenuated profile. However, this adaptive subtraction method highly depends on the initial prediction of the ground-roll noise. Besides, in the case of highly non-stationary primary reflections and ground-roll noise, a conventional stationary matching filtering is not physically reasonable, and thus can not obtain a satisfactory performance. In the next sections, we will introduce a novel approach for attenuating ground-roll noise based on bandpass filtering, which can remove more ground-roll noise and preserve more useful primary reflections.

\subsection{Local signal-and-noise orthogonalization}
%\cite{yangkang2014ortho} proposed a local orthogonalization approach for retrieving useful signals from the noise section. 
The principle of local signal-and-noise orthogonalization is to locally orthogonalize the denoised signal and noise sections in order to make sure no coherent primary reflections will be lost in the noise section. %\cite{yangkang2014ortho} mainly dealt with random noise attenuation problem.
Here, we propose to use this algorithm to ground-roll noise removal problem. The orthogonalization based approach can be summarized as: 
\begin{align}
\label{eq:news}
\mathbf{s}&=(\mathbf{I}+\mathbf{W})\mathbf{\hat{d}},\\
\label{eq:newn}
\mathbf{n}&=\mathbf{\hat{d}}-(\mathbf{I}+\mathbf{W})\mathbf{\hat{d}}.
\end{align}
Here, $\mathbf{s}$ and $\mathbf{n}$ are the output denoised signal and noise sections.\wen{$\mathbf{I}$ denotes an identity matrix.} $\mathbf{W}$ denotes a diagonal operator composed of the local orthogonalization weight (LOW) vector. \dlo{The LOW closely follows the way calculating the global orthogonalization weight (GOW).} When $w=\frac{\mathbf{n}_0^T\mathbf{s}_0}{\mathbf{s}_0^T\mathbf{s}_0}$, \wen{$\mathbf{s}_0$ and $\mathbf{n}_0$ denoting the initial guess of signal and noise} , $w$ is the global orthogonalization weight (GOW), it can be proved that the following scaled signal $\mathbf{s}_0$ and corresponding noise $\mathbf{n}_0$ are orthogonal to each other in a global sense: %\cite{yangkang2014ortho}:
%\begin{figure}[htb!]
%  \centering
%  \includegraphics[width=0.32\columnwidth,height=0.32\columnwidth]{Fig/demo}
%  \caption{Demonsration of global orthogonalization.}
%   \label{fig:demo}
%\end{figure}

\begin{align}
\label{eq:ortho1}
\hat{\mathbf{n}} &= \mathbf{n}_0 - w\mathbf{s}_0, \\
\label{eq:ortho2}
\hat{\mathbf{s}} &= \mathbf{s}_0 + w\mathbf{s}_0.
\end{align}
In a local sense, the LOW can be define as:
      \begin{equation}
        \label{eq:eq2}
          w_m(t) = \frac{\displaystyle\sum_{i=t-m/2}^{t+m/2} s_0(i) n_0(i)}{\displaystyle\sum_{i=t-m/2}^{t+m/2} s_0^2(i)},
      \end{equation}
where $w_m(t)$ denotes the LOW for each temporal point $t$ with a local window length $m$. $s_0(t)$ and $n_0(t)$ here denote the initially estimated signal and noise for each point $t$.

In order to better control the locality and smoothness of LOW, we follow the local-attribute scheme introduced by \cite{fomel2007localattr}: %and first turn the orthogonalization problem into an optimization problem:
\begin{equation}
\label{eq:localortho}
\mathbf{w} = \arg\min_{\mathbf{\tilde{w}}} \parallel \mathbf{n}_0 - \mathbf{S}_0\mathbf{\tilde{w}}\parallel_2^2. %+ \mathbf{R}(\mathbf{w})
\end{equation}
Here, $\mathbf{w}$ is the LOW, $\mathbf{S}_0$ is a diagonal matrix composed of the initial estimated signal $\mathbf{s}_0$: $\mathbf{S}_0=diag(\mathbf{s}_0)$.
Then, we solve the least-squares problem \ref{eq:localortho} with the help of shaping regularization using a local-smoothness constraint:
\begin{equation}
\label{eq:shape}
\mathbf{w} = [\lambda^2\mathbf{I} + \mathcal{T}(\mathbf{S}_0^T\mathbf{S}_0-\lambda^2\mathbf{I})]^{-1}\mathcal{T}\mathbf{S}_0^T\mathbf{n}_0,
\end{equation}
where $\mathcal{T}$ is a triangle smoothing operator and $\lambda$ is a scaling parameter set as $\lambda  = \Arrowvert\mathbf{S}_0^T\mathbf{S}_0\Arrowvert_2$ \cite{fomel2007localattr}. The triangle smoothing operator was introduced in detailed in \cite{fomel2007shape}. It should be mentioned that solution of equation \ref{eq:localortho} corresponds to a regularized division between the two vectors $\mathbf{n}_0$ and $\mathbf{s}_0$ and it can be solved using any regularization approach, not limited to the shaping regularization strategy shown in equation \ref{eq:shape}. Thus it is fairly convenient to implement the local orthogonalization between initial signal and noise.

\subsection{Local bandlimited orthogonalization}
The basic principle of the proposed approach is to first apply a simple bandpass filtering to the raw common shot gather in order to obtain an initial guess of primary reflections and ground-roll noise. Then we orthogonalize the initial guess of primary reflections and ground-roll noise to obtain a signal compensated primary reflection gather, and a signal-free ground-roll noise section. In order to make the local orthogonalization robust, we need first use a relative high LBF such that all the ground-roll noise is removed initially. This proposed workflow for attenuating ground-roll noise is termed as the \emph{local bandlimited orthogonalization}. The detailed workflow of the proposed approach is shown in Figure \ref{fig:flowchart}. As we can see, the total workflow of the proposed approach is fairly convenient to implement, because there are only two steps involved in the processing. First, using a conventional bandpass filtering. Second, using local signal-and-noise orthogonalization. As we stated in the last section, the local signal-and-noise orthogonalization is no more than solving a regularized division problem (equations \ref{eq:eq2} and \ref{eq:localortho}). Thus, the proposed approach can be conveniently used as an alternative in the industry. In the next section, we will use a typical dataset for demonstrating the effective performance of the proposed approach.
\begin{figure}[htb!]
  \centering
    \includegraphics[width=0.8\columnwidth]{Fig/flowchart}
	\caption{Workflow of the proposed ground-roll noise attenuation approach based on local bandlimited orthogonalization.}
   \label{fig:flowchart}
\end{figure}
\section{Example}
The dataset that we use to demonstrate the effectiveness of our proposed denoising approach is \wen{an open-source dataset, which has been widely used in the geophysics community \cite{yilmaz1987,carson2006}.} \dlo{borrowed from  and }\wen{This data }is referred to as the OZ-25 dataset \cite{carson2006}. \wen{One can easily download it either from the Seismic Unix (SU) website ($http://www.cwp.mines.edu/cwpcodes/$) or from the Madagascar software website ($www.ahay.org$, \cite{mada2013}).} The raw data in common shot domain is shown in Figure \ref{fig:field}. \wen{The temporal sampling rate is 0.002s and the spatial sampling rate is 0.05km. The shot location is located at the middle of the receiver array, thus the shot record has a symmetric form. There are 81 traces (receivers) in the shot record and the offset ranges from -2km to 2km.} The OZ-25 dataset is a typical dataset suitable for testing the ground-roll noise attenuation performance because of several reasons. First,  there is a large amount of ground-roll noise contaminating this dataset. It is obvious that most of the primary reflections are covered by this coherent ground-roll noise. Secondly, the reflections are highly non-stationary. We can clearly see that the amplitude of the reflections are variable in the range of the whole gather. Because of the large amplitude range of primary reflections, it can be effectively used to test the primary reflections preservation properties of a certain denoising algorithm.

\dlo{Figures \ref{fig:field-ortho-0} and \ref{fig:dif-ortho-0} show the denoised section and noise section using the proposed approach.} We locally orthogonalize the denoised section and noise section that come from the bandpass filtering using $fl=25$ Hz. The initial guess of the primary reflections and ground-roll noise are shown in Figures \ref{fig:field-1-0} and \ref{fig:dif-1-0}. \wen{Figures \ref{fig:field-ortho-0} and \ref{fig:dif-ortho-0} show the denoised section and noise section using the proposed approach.} The denoised data using the proposed approach shows excellent result, because most of the ground-roll noise have been removed but no primary reflections energy is damaged.

\wen{We also apply the widely used adaptive subtraction method to the field data and show its performance in Figures \ref{fig:field-3-0} and \ref{fig:dif-3-0}. Figures \ref{fig:field-3-0} and \ref{fig:dif-3-0} correspond to the denoised data and removed noise, respectively. It is obvious that there is some residual ground-roll noise existing in the denoised data and the noise section contains significant useful reflections, especially for the shallow part.}

We zoom several parts from the denoised sections and noise sections using different approaches and show them in Figures \ref{fig:zooma-1,zooma-2,zooma-ortho,zooma-3} and \ref{fig:zoomb-1,zoomb-2,zoomb-ortho,zoomb-3}. Figure \ref{fig:zooma-1,zooma-2,zooma-ortho,zooma-3} shows the zoomed denoised sections for the frame box A, as shown in Figures \ref{fig:field-1-0},\ref{fig:field-2-0}\dlo{ and }\wen{, }\ref{fig:field-ortho-0}, \wen{and \ref{fig:field-3-0},} respectively. We can see there is an obvious primary reflections recovery using the proposed approach by comparing Figures \ref{fig:zooma-1} and \ref{fig:zooma-ortho}. There is also an obvious decrease of noise using the proposed approach by comparing Figures \ref{fig:zooma-2} and \ref{fig:zooma-ortho}. \wen{It is more obvious that the adaptive subtraction method causes some residual dipping ground-roll noise energy by comparing Figures \ref{fig:zooma-ortho} and \ref{fig:zooma-3}.}

Figure \ref{fig:zoomb-1,zoomb-2,zoomb-ortho,zoomb-3} shows the zoomed noise sections for frame box B, as shown in Figures \ref{fig:dif-1-0},\ref{fig:dif-2-0}\dlo{ and}\wen{, } \ref{fig:dif-ortho-0},\wen{and \ref{fig:dif-3-0},} respectively. Please note that there is a decrease of primary reflections using the proposed approach, comparing Figures \ref{fig:zoomb-1} and \ref{fig:zoomb-ortho}, and there is an increase of noise removal using the proposed approach, comparing Figures \ref{fig:zoomb-2} and \ref{fig:zoomb-ortho}. \wen{The proposed approach and the adaptive subtraction method are very similar in this selected region. There is a slight amount of useful energy in Figure \ref{fig:zoomb-ortho} but hopefully not significant, considering the overall denoising performance. The ground-roll noise energy in Figure \ref{fig:zoomb-ortho}, however, is a bit stronger than that in Figure \ref{fig:zoomb-3}.} These observations indicate that compared with bandpass filtering with $fl=25$ Hz, the proposed approach preserves much more useful energy, and compared with bandpass filtering with $fl=10$ Hz, the proposed method removes much more ground-roll noise. \wen{Compared with the traditional adaptive subtraction method, the proposed approach can obtain a generally better performance}.

Figure \ref{fig:field-fs} shows a comparison of the average spectrum of all the traces for different data. \dlo{The black solid line denotes the average spectrum of raw data. The green dash line corresponds to the proposed approach. The red dot dash line corresponds to $fl=25$ Hz. The yellow long dash line corresponds to $fl=10$ Hz. We can see obviously that there is a removal of ground-roll noise spectrum from the black solid and yellow long dash lines to the green dash line. There is also a spectrum boost of the primary reflections from the red dot dash line to green dash line.}\wen{The black solid line denotes the average spectrum of raw data. The green dashed line corresponds to the proposed approach. The red dot-dashed line corresponds to the high-pass filtering with $fl=25$ Hz. The yellow long-dashed line corresponds to the high-pass filtering with $fl=10$ Hz. The blue dotted line corresponds to the adaptive subtraction method. We can see obviously that there is a removal of ground-roll noise spectrum from the black solid and the yellow long-dashed lines to the green dashed line.  There is also a spectrum boost of the primary reflections from the red dot-dashed line to the green dashed line. There exists several spectrum notches in the blue dotted line, which indicates that there is still a mixture of useful reflections and ground-roll noise after applying the adaptive subtraction method.}

\section{Conclusions}
We have proposed a novel local bandlimited  orthogonalization approach for removing highly non-stationary ground-roll noise, which can remove most of ground-roll noise without harming the useful primary reflections. We orthogonalize the initial guess of primary reflections and ground-roll noise using local signal-and-noise orthogonalization. The initial guess of primary reflections and ground-roll noise are bandlimited data from a common bandpass filtering with a relatively high LBF such that all the ground-roll noise is removed during the initial guess. The proposed approach can solve the frequency-mixing problem when applying a simple bandpass filtering. The proposed approach can guarantee that the least amount of useful primary reflections is lost in the noise section. The procedure of the proposed approach is fairly convenient to implement because only a bandpass filtering and a regularized division between the initially denoised signal and initial noise are used. We have used an open-source field dataset to demonstrate the successful performance of the proposed approach in real data processing.

\begin{figure}[htb!]
  \centering
    \includegraphics[width=.42\columnwidth]{field/Fig/field}
	\caption{Raw OZ-25 field data, borrowed from \cite{carson2006}.}
   \label{fig:field}
\end{figure}

\begin{figure}[htb!]
  \centering
    \subfigure[]{\includegraphics[width=0.41\columnwidth]{field/Fig/field-1-0}
    \label{fig:field-1-0}}
    \subfigure[]{\includegraphics[width=0.41\columnwidth]{field/Fig/dif-1-0}
    \label{fig:dif-1-0}}   
    \subfigure[]{\includegraphics[width=0.41\columnwidth]{field/Fig/field-2-0}
    \label{fig:field-2-0}}
    \subfigure[]{\includegraphics[width=0.41\columnwidth]{field/Fig/dif-2-0}
    \label{fig:dif-2-0}}         
	\caption{(a) Bandpass filtered data (fl=25 Hz). (b) Difference section corresponding to (a). (c) Bandpass filtered data (fl=10 Hz). (d) Difference section corresponding to (c).}
   \label{fig:field-1-0,dif-1-0,field-2-0,dif-2-0}
\end{figure}

\begin{figure}[htb!]
  \centering
    \subfigure[]{\includegraphics[width=0.41\columnwidth]{field/Fig/field-ortho-0}
    \label{fig:field-ortho-0}}
    \subfigure[]{\includegraphics[width=0.41\columnwidth]{field/Fig/dif-ortho-0}
    \label{fig:dif-ortho-0}}   
    \subfigure[]{\includegraphics[width=0.41\columnwidth]{field/Fig/field-3-0}
    \label{fig:field-3-0}}
    \subfigure[]{\includegraphics[width=0.41\columnwidth]{field/Fig/dif-3-0}
    \label{fig:dif-3-0}}         
	\caption{(a) Denoised data using the proposed approach. (b) Noise section corresponding to (a). (c) Denoised data using the adaptive subtraction approach. (d) Noise section corresponding to (c).}
   \label{fig:field-ortho-0,dif-ortho-0,field-3-0,dif-3-0}
\end{figure}

\begin{figure}[htb!]
  \centering
    \subfigure[]{\includegraphics[width=0.38\columnwidth]{field/Fig/zooma-1}
    \label{fig:zooma-1}}
    \subfigure[]{\includegraphics[width=0.38\columnwidth]{field/Fig/zooma-2}
    \label{fig:zooma-2}}   
    \subfigure[]{\includegraphics[width=0.38\columnwidth]{field/Fig/zooma-ortho}
    \label{fig:zooma-ortho}}
    \subfigure[]{\includegraphics[width=0.38\columnwidth]{field/Fig/zooma-3}
    \label{fig:zooma-3}}         
	\caption{(a)-(d) Zoomed denoised section comparisons for frame box A (as shown in Figures \ref{fig:field-1-0},\ref{fig:field-2-0} , \ref{fig:field-ortho-0}, and \ref{fig:field-3-0}, respectively). Note the primary reflections recovery from (a) to (c) and the noise decrease from (b) to (c). There is obvious residual ground-roll noise existing in (d).}
   \label{fig:zooma-1,zooma-2,zooma-ortho,zooma-3}
\end{figure}

\begin{figure}[htb!]
  \centering
    \subfigure[]{\includegraphics[width=0.38\columnwidth]{field/Fig/zoomb-1}
    \label{fig:zoomb-1}}
    \subfigure[]{\includegraphics[width=0.38\columnwidth]{field/Fig/zoomb-2}
    \label{fig:zoomb-2}}   
    \subfigure[]{\includegraphics[width=0.38\columnwidth]{field/Fig/zoomb-ortho}
    \label{fig:zoomb-ortho}}
    \subfigure[]{\includegraphics[width=0.38\columnwidth]{field/Fig/zoomb-3}
    \label{fig:zoomb-3}}         
	\caption{(a)-(d) Zoomed noise section comparisons for frame box B (as shown in Figures \ref{fig:dif-1-0},\ref{fig:dif-2-0}, \ref{fig:dif-ortho-0}, and \ref{fig:dif-3-0}, respectively). Note the decrease of primary reflections from (a) to (c) and the increase of noise removal from (b) to (c). (c) and (d) are very similar in this zoomed region.}
   \label{fig:zoomb-1,zoomb-2,zoomb-ortho,zoomb-3}
\end{figure}

\begin{figure}[htb!]
  \centering
    \includegraphics[width=0.9\columnwidth]{field/Fig/field-fs}
	\caption{Comparisons of the average spectrum of all the traces. \dlo{Black solid line denotes the average spectrum of raw data. Green dash line corresponds to the proposed approach. Red dot dash line corresponds to $fl=25$ Hz. Yellow long dash line corresponds to $fl=10$ Hz. Note the removal of ground-roll noise spectrum from the black solid and yellow long dash lines to the green dash line, and the primary reflections spectrum boost from the red dot dash line to green dash line.}\wen{The black solid line denotes the average spectrum of raw data. The green dashed line corresponds to the proposed approach. The red dot-dashed line corresponds to $fl=25$ Hz. The yellow long-dashed line corresponds to $fl=10$ Hz. The blue dotted line corresponds to the adaptive subtraction method. Note the removal of ground-roll noise spectrum from the black solid and the yellow long-dashed lines to the green dashed line, and the primary reflections spectrum boost from the red dot-dashed line to the green dashed line. There exists several spectrum notches in the blue dotted line, indicating a mixture of useful reflections and ground-roll noise after applying the adaptive subtraction method.} }
   \label{fig:field-fs}
\end{figure}

\section{Acknowledgments}
We would like to thank Qun Luo, Shan Qu, Jiang Yuan for constructive suggestions \wen{that improved the manuscript greatly}. Yangkang Chen appreciate\wen{s} Sergey Fomel for the inspiring discussions about local orthogonalization. \wen{The paper is reproducible within the Madagascar open-source platform \cite{mada2013}.} We are grateful to developers of the Madagascar software package for providing corresponding codes for testing the algorithms and preparing the figures. \wen{This work is partially supported by the Natural Science Fund of China (grant nos.:91330108, 41374121, and 61327013), the fundamental Research Funds for the Central Universities (grant no.: HIT.BRETIV.201314), the Program for New Century Excellent Talents in University (grant no.: NCET-11-0804), the China Postdoctoral Science Foundation (grant no.: 2014M561053), the Natural Science Foundation of Hebei Province (grant no.: F2013202254), and the Texas Consortium for Computational Seismology (TCCS).}

\bibliographystyle{IEEEtran}
\bibliography{groll}