### 2021-01-12 rtseis/paper.tex
up

\title{Relative-time seislet transform}
\author{Zhicheng Geng~\footnotemark[1], 
        Xinming Wu~\footnotemark[1],  
        Sergey Fomel~\footnotemark[1] and
        Yangkang Chen~\footnotemark[2]}

\address{
    \footnotemark[1]Bureau of Economic Geology \\
    John A. and Katherine G. Jackson School of Geosciences \\
    The University of Texas at Austin \\
    University Station, Box X \\
    Austin, Texas, USA \\
    \footnotemark[2]Zhejiang University\\
    School of Earth Sciences\\
    Hangzhou, Zhejiang, China
}

\righthead{RT-seislet transform}
\lefthead{Geng et al. (2019)}

\maketitle

\begin{abstract}
    The seislet transform utilizes the wavelet-lifting scheme and local slopes 
    to analyze the seismic data. 
    In its definition, the designing of prediction operators specifically for 
    seismic images and data is an important issue. 
    We propose a new formulation of the seislet transform based on the relative time (RT) attribute. 
    This method uses RT volume to construct multiscale prediction 
    operators. With the new prediction operators, the seislet transform gets accelerated 
    since distant traces get predicted directly. 
    We apply the proposed method to synthetic and real data to demonstrate that 
    the new approach reduces computational cost and obtains excellent sparse
    representation on test datasets. %\par
\end{abstract}

%-----------------------------------------------------------------
\section{Introduction}
    Sparse transforms aim to represent the most important information of an 
    image with few coefficients in the transform domain while obtaining a good 
    quality approximation of the original image. 
    Over the past several decades, different types of wavelet-like transforms 
    have been proposed and successfully applied in image compression and 
    denoising~\cite[]{Hennen2006, Li2013}, including 
    curvelets~\cite[]{Starck2002, Ma2010}, contourlets~\cite[]{Do2005}, 
    shearlets~\cite[]{Demetrio2005}, directionlets~\cite[]{Velisavljevic2006}, 
    bandelets~\cite[]{LePennec2005}. 
    The strong anisotropic selectivity of these wavelet-like transforms helps 
    achieve excellent data compression and accurate reconstruction for seismic 
    images.

    \cite{Fomel2010a} introduced the seislet transform, which is a digital 
    wavelet-like transform designed specifically for seismic data. 
    Based on the lifting scheme used in digital wavelet transform (DWT) 
    construction~\cite[]{Sweldens1995}, the seislet transform follows dominant 
    local slopes obtained by plane-wave destruction (PWD)~\cite[]{Claerbout1992,
    Fomel2002} to predict seismic events. 
    Instead of using PWD, \cite{Liu2010} used offset continuation (OC) to 
    construct updating and prediction operators. 
    The OC-seislet transform has better performance than the PWD-seislet 
    transform in characterizing and compressing structurally complex prestack 
    reflection data. 
    To reduce sensitivity to strong noise interference, \cite{Liu2015} proposed 
    a velocity-dependent (VD) seislet transform. 
    In this method, the normal moveout equation was introduced to serve as a 
    connection between local slope and scanned velocities. 
    \cite{Chen2018} utilized empirical mode decomposition (EMD) to obtain 
    smoothly non-stationary data, and then applied the 1D non-stationary seislet 
    transform to the data. This new approach was called the EMD-seislet 
    transform and has shown excellent performance in attenuating random noise. 
    Recently, several studies show the superiority of the seislet transform
    on sparse representation of seismic data over the Fourier transform, the 
    wavelet transform and the curvelet transform \cite[]{Fomel2010a, Gan2015, 
    Chen2018}.

    With the ability to compress and reconstruct seismic images, the 
    seislet transform has been successfully applied in seismic data processing 
    such as noise attenuation \cite[]{Chen2016a, Chen2016b}, deblending 
    \cite[]{Chen2014, Gan2016} and data interpolation \cite[]{Cai2013, shuwei2016cs}, 
    However, the original implementation of PWD-seislet transform requires
    recursively computation when predicting distant traces, which increases its 
    computation cost, especially when the dataset is very large.
    Additionally, the smooth slopes from PWD in the original seislet transform 
    can fail to correctly follow reflections across discontinuities including 
    faults and unconformities. 
    Therefore, geologically meaningful discontinuities may not be optimally 
    compressed.

    The \old{Relative Time}\new{relative time} (RT) \old{Volume}\new{volume}, $\tau(x,y,t)$, has different meanings in
    different domains. 
    In seismic image domain, the \old{Relative Time}\new{RT} is the same as the 
    \old{Relative Geologic Time}\new{relative geologic time}~\cite[]{Stark2003, Stark2004} and each RT contour 
    corresponds to a geologic horizon.
    However, RT volumes of seismic gathers \old{don't}\new{do not} necessarily have geologic 
    meaning, since in this case, constant RT contours align with seismic events.
    There are several ways to generate an RT volume. 
    One can always pick out as many horizons or events as possible to obtain it, 
    which is simple but laborious. 
    The RT volume can also be generated by unwrapping seismic instantaneous 
    phase~\cite[]{Stark2003, Wu2012}.
    \cite{Fomel2010b} used local slopes of seismic events estimated by 
    PWD~\cite[]{Fomel2002} to generate the RT volume, which has superior 
    computational performance. 
    The RT volume already has \old{successful applications}\new{been successfully applied} in missing well-log 
    data prediction \cite[]{Bader2018}\new{ and} seismic horizons construction 
    \cite[]{Wu2015}\old{ and sedimentological interpretation}.

    In this paper, we propose a new implementation of the seislet transform, 
    called RT-seislet transform, by using the \old{Relative Time}\new{RT} 
    obtained from the predictive painting ~\cite[]{Fomel2010b} to construct 
    prediction operators for the seislet transform. 
    In this way, prediction of one trace from distant traces is computed 
    directly and accurately, which saves a lot of computation and has better 
    performance with faults and unconformities.
    By applying the new formulation of the seislet transform to synthetic and 
    real data, we demonstrate that the proposed method is efficient and able to 
    preserve discontinuities, while achieving excellent sparse representation of 
    seismic data.

    This paper is organized as follows: we first review the seislet 
    transform and the estimation of \new{a} RT volume. 
    Then, we incorporate RT attribute to construct new formulation of prediction 
    and update operators for the seislet transform.
    Finally, we test the proposed RT-seislet transform on both synthetic and 
    real datasets and compare its performance with the PWD-seislet transform.

%-----------------------------------------------------------------
\section{Theory}
\subsection{Reviews of 2D Seislet Transform}
    As proposed by \cite{Fomel2010a}, the seislet transform can be constructed 
    using the lifting scheme \cite[]{Sweldens1995}. 
    The procedure of the lifting scheme is as follows:
    \begin{enumerate}
        \item Arrange the input data as a sequence of records. 
            In the seislet transform, the input data \old{is}\new{are} common shot gather or 
            seismic image.
        \item Divide the arranged records into even and odd components 
            $\mathbf{e}$ and $\mathbf{o}$. For the 2D seislet transform, each 
            trace is a component and the seismic data \old{is}\new{are} split into two parts 
            according to trace indices.
        \item Find the residual $\mathbf{r}$ between the odd component and 
            its prediction from the even component
            \begin{equation}
                \mathbf{r} = \mathbf{o} - \mathbf{P[e]},
                \label{eq:a}
            \end{equation}
            where $\mathbf{P}$ is a \emph{prediction} operator.
        \item Using the difference from the previous step to get a coarse 
            approximation $\mathbf{c}$ of the data by updating the even 
            component
            \begin{equation}
                \mathbf{c} = \mathbf{e} + \mathbf{U[r]},
                \label{eq:b}
            \end{equation}
            where $\mathbf{U}$ is an \emph{update} operator.
        \item The coarse approximation $\mathbf{c}$ obtained from last step 
            becomes the new data, and previous steps are applied to the new data 
            at the next scale level.
    \end{enumerate}

    The coarse approximation $\mathbf{c}$ from the final scale and residual
    difference $\mathbf{r}$ from all scales form the result of the seislet 
    transform.

    The prediction and update operators are defined, for example, by modifying 
    operators in the construction of CDF 5/3 biorthogonal 
    wavelets~\cite[]{Cohen1992} as follows:
    \begin{equation}
        \label{P}
        \mathbf{P[e]}_k = \left(\mathbf{S}_k^{(+)}[\mathbf{e}_{k-1}]+
        \mathbf{S}_k^{(-)}[\mathbf{e}_{k}] \right) / 2
    \end{equation}
    and
    \begin{equation}
        \label{U}
        \mathbf{U[r]}_k = \left(\mathbf{S}_k^{(+)}[\mathbf{r}_{k-1}]+
        \mathbf{S}_k^{(-)}[\mathbf{r}_{k}] \right) / 4,
    \end{equation}
    where $\mathbf{S}_k^{(+)}$ and $\mathbf{S}_k^{(-)}$ can accomplish the 
    prediction of a trace from its left and right neighbors, respectively. 
    The predictions need to be applied at different scales. 
    For 2D seislet transform, different scales mean different distances 
    between traces. 
    More accurate higher-order formulations are possible. 
    Following the inverse lifting scheme from coarse \old{scale }to fine scale, 
    the inverse seislet transform can be constructed.

\subsection{Relative Time volume estimation}
    In the proposed method, \old{the }predictive painting~\cite[]{Fomel2010b}, \old{which
    is }based on the prediction operator arising from PWD~\cite[]{Fomel2002}, is 
    utilized to generate the \old{Relative Time}\new{RT} volume\old{,}\new{.} 
    In the predictive painting, after local slopes are estimated, an RT 
    volume is computed by spreading time values of reference traces along 
    dominant event slopes throughout the whole seismic volume.

    Using linear operator notation, PWD operation is defined as
    \begin{equation}
        \mathbf{d} = \mathbf{D}{}\mathbf{s},
    \end{equation}
    where $\mathbf{s} =[\mathbf{s}_1 \,\mathbf{s}_2 \,\cdots\, \mathbf{s}_N]^T$ 
    is a collection of seismic traces and $\mathbf{d}$ denotes the residual from 
    destruction. $\mathbf{D}$ is the destruction operator defined as 
    \begin{equation}
        \label{pred}
        \mathbf{D}=
        \begin{bmatrix}
            \mathbf{I} & 0 & 0 & \cdots & 0 \\
            -\mathbf{P}_{1,2} & \mathbf{I} & 0 & \cdots & 0 \\
            0 & -\mathbf{P}_{2,3} & \mathbf{I} & \cdots & 0 \\
            \vdots & \ddots & \ddots & \ddots & \vdots \\
            0 & \cdots & 0 & -\mathbf{P}_{N-1,N} & \mathbf{I}
        \end{bmatrix},
    \end{equation}
    where $\mathbf{I}$ denotes the identity matrix and $\mathbf{P}_{i,i+1}$ is 
    an operator that predicts the $(i+1)$-th trace from the $i$-th trace 
    according to the local slope of seismic event or seismic horizon.

    By a simple recursion, a trace can be predicted utilizing a distant trace. 
    For example, $\mathbf{P}_{r,k}$, predicting the $k$-th trace from the $r$-th 
    trace, is \old{simply }defined as
    \begin{equation}
        \label{predict}
        \mathbf{P}_{r,k}=\mathbf{P}_{k-1,k}\mathbf{P}_{k-2,k-1}\cdots
        \mathbf{P}_{r+1,r+2}\mathbf{P}_{r,r+1}.
    \end{equation}
    Then, the prediction of trace $\mathbf{s}_r$, given a reference trace
    $\mathbf{s}_r$ is accomplished by
    \begin{equation}
        \label{pp}
        \mathbf{s}_k^{\text{pre}}=\mathbf{P}_{r,k}\mathbf{s}_r.
    \end{equation}
    After determining the prediction operators in 
    \old{E}\new{e}quation~\ref{pred} by PWD, 
    the predictive painting algorithm recursively spreads information from a 
    reference trace to the whole seismic data volume by following local slopes
    using \new{e}\old{E}quation \ref{pp}. 
    If the information spread by the predictive painting \old{is}\new{are} the time \old{value}\new{coordinates}, an 
    RT volume is easily computed.

    Figure~\ref{fig:sigmoid} shows a synthetic image with several complex 
    geologic structures. 
    Local slopes corresponding to Figure~\ref{fig:sigmoid} are shown in 
    Figure~\ref{fig:dip-pad}, which \old{is}\new{are} estimated by PWD. 
    Figure~\ref{fig:pick-1} is the \old{Relative Time}\new{RT} volume calculated 
    by the predictive painting using local slopes in Figure~\ref{fig:dip-pad}. 
    The reference trace of this RT volume is in the center of the image, which 
    is marked as a vertical line in Figure~\ref{fig:pick-1}. 
    This single reference RT volume is used to flatten the input seismic 
    image (Figure \ref{fig:sigmoid}) by unshifting each traces using the 
    information it contains, since \old{Relative Time}\new{RT} indicates how 
    much a given trace is shifted with respect to the reference trace.
    The flattened image is shown in Figure~\ref{fig:flat-1}.
    Another RT volume with multiple references, shown in 
    Figure~\ref{fig:pick-5}, is generated by selecting five reference traces and 
    calculating the average of all estimated \old{Relative Time}\new{RT} 
    volumes.
    Figure~\ref{fig:flat-5} shows the corresponding flattened image.
    The comparison between two flattened seismic images (Figure~\ref{fig:flat-1} 
    and \ref{fig:flat-5}) indicates that the RT volume with only one reference 
    trace \old{doesn't}\new{does not} necessarily have information about all seismic events since 
    there are discontinuities such as faults and unconformities, while the RT 
    volume computed with more reference traces contains more information.

    \inputdir{rt}
    \plot{sigmoid}{width=0.8\columnwidth}{Synthetic seismic image from 
        \cite{Claerbout2000}.}
    \plot{dip-pad}{width=0.8\columnwidth}{Local slopes estimated from 
        Figure~\ref{fig:sigmoid}.}
    \multiplot{2}{pick-1,pick-5}{width=0.6\columnwidth}{\old{Relative time}
        \new{RT} estimated by predictive painting from Figure~\ref{fig:sigmoid} 
        using (a) a single reference trace and (b) multiple reference traces.}
    \multiplot{2}{flat-1,flat-5}{width=0.6\columnwidth}{Flattened images using
        RT volumes from (a) Figure \ref{fig:pick-1} and (b) 
        Figure~\ref{fig:pick-5}.}

\subsection{RT formulation of prediction and update operators}
    \old{Relative time}\new{RT} contains the relationship between traces. 
    From RT volume, it is straightforward to obtain the shift of one trace with 
    respect to the reference. 
    Therefore, the trace prediction using RT volumes is formulated as follows:
    \begin{equation}
        \mathbf{s}_k^{\text{pre}}(t)=\mathbf{s}_r(\mathbf{\tau}_{k,r}(t))
        \label{new_form}
    \end{equation}
    where $\mathbf{s}_k$ and $\mathbf{s}_r$ denote the $k$-th and the $r$-th 
    trace, respectively. 
    $\tau_{k,r}(t)$ is the RT value of the $k$-th trace with the $r$-th trace as 
    the reference trace, indicating the shift of the $k$-th trace with
    respect to the $r$-th trace. 
    According to \new{e}\old{E}quation \ref{new_form}, the prediction of the 
    $k$-th trace from the $r$-th trace is easily accomplished by simply applying 
    forward interpolation to the $r$-th trace using corresponding RT values.
    In the proposed method, we use \new{e}\old{E}quation \ref{new_form} to 
    define operators $\mathbf{S}$ in \new{e}\old{E}quations \ref{P} and 
    \ref{U} to construct corresponding prediction and update operators. 
    In this way, a trace is predicted from a distant trace directly by the 
    \old{Relative Time}\new{RT} attribute instead of the recursive computation 
    used in the PWD-seislet transform. 
    Also, because an accurate RT volume stores information about all horizons 
    and structural discontinuities including faults and unconformities, 
    predictions of traces around discontinuities using \new{e}\old{E}quation 
    \ref{new_form} are accurate. 
    Therefore, a better \old{approximation}\new{delineation} of faults and unconformities can be 
    achieved.

    In computing an RT volume, we need to first choose one or multiple reference 
    traces. 
    After computing the RT volume, the \old{Relative Time}\new{RT} attribute 
    between any two traces is easily obtained from one RT volume using\old{ the 
    following formulation}:
    \begin{equation}
        \label{rt}
        \mathbf{\tau}_{k,r}(t)=\mathbf{\tau}^{-1}_{r,l}(\mathbf{\tau}_{k,l}(t)),
    \end{equation}
    where $\mathbf{\tau}_{j,i}(t)$ represents the RT value of the $j$-th trace 
    with the $i$-th trace as the reference trace, and 
    $\mathbf{\tau}^{-1}_{r,l}(t)$ is the time warping \cite[]{Burnett2009} of 
    $\tau_{r,l}(t)$. 
    According to \new{e}\old{E}quation \ref{rt}, given the shift information 
    $\tau_{r,l}(t)$ and $\tau_{k,l}(t)$, the shift relationship between the 
    $k$-th trace and the $r$-th trace can be obtained by applying an inverse 
    interpolation to $\tau_{r,l}(t)$ using $\tau_{k,l}(t)$. 
    Then, \new{e}\old{E}quation \ref{new_form} is applied to implement the 
    prediction of the $k$-th trace from the $r$-th trace.
    Therefore, only one RT volume is needed for the proposed implementation of 
    the seislet transform.

    Figure~\ref{fig:pick-150,invint,pick-50,pick-50-true} shows an example of
    \old{how to utilize}\new{implementing} \new{e}\old{E}quation~\ref{rt} to compute an RT volume from 
    another \old{RT volume}\new{one} with different reference traces. 
    The RT volume in Figure~\ref{fig:pick-150} is obtained by the predictive 
    painting with the 150th trace as the reference trace. 
    This RT volume can be represented by $\mathbf{\tau}_{k,150}(t), 
    k=1,2,\ldots,n$. 
    Figure~\ref{fig:invint} is the time warping of this RT volume. 
    We extract the 50th trace from Figure~\ref{fig:invint}, which is denoted by 
    $\mathbf{\tau}^{-1}_{50,150}(t)$. 
    Then, by implementing \new{e}\old{E}quation~\ref{rt}, i.e., inverse 
    interpolating $\mathbf{\tau}^{-1}_{50,150}(t)$ using the whole RT volume 
    ($\mathbf{\tau}_{k,150}(t), k=1,2,\ldots,n$), we can get a new 
    RT volume, \new{as }shown in Figure~\ref{fig:pick-50} ($\mathbf{\tau}_{k,50}(t), 
    k=1,2,\ldots,n$). Here, we use the 50th trace as the reference trace. 
    To evaluate the effectiveness of \new{e}\old{E}quation \ref{rt}, RT 
    volumes from Figure \ref{fig:pick-50} and \ref{fig:pick-50-true} are used to 
    flatten Figure \ref{fig:sigmoid}. 
    Flattened images are shown in Figure \ref{fig:flat1,flat2}. Small 
    differences between Figure \ref{fig:flat1} and \ref{fig:flat2} indicate that 
    these two RT volumes (Figure \ref{fig:pick-50} and \ref{fig:pick-50-true}) 
    contain similar information, which \old{proves}\new{demonstrate} that 
    \new{e}\old{E}quation~\ref{rt} can be used to get the relationship between 
    any two traces from one single RT volume.

    \multiplot{4}{pick-150,invint,pick-50,pick-50-true}{width=0.4\columnwidth}{
        (a) The RT volume obtained by the predictive painting and the reference 
        trace is the 150th trace. (b) Time warping of (a). (c) RT 
        volume obtained by \new{e}\old{E}quation~\ref{rt}. (d) The RT volume 
        obtained by the predictive painting and the reference trace is the 50th 
        trace.}
    \multiplot{3}{flat1,flat2}{width=0.7\columnwidth}{Flattened images using RT 
        volumes from (a) Figure \ref{fig:pick-50} and (b) 
        Figure \ref{fig:pick-50-true}.}

\subsection{Workflow}
    The workflow of the proposed method is shown in Figure~\ref{fig:workflow}. 
    As we can see, the proposed approach is convenient to implement since only 
    one RT volume is needed before applying the new formulation of the seislet 
    transform. 
    In the proposed approach, we need first use PWD to compute the local slopes 
    of the input seismic data such that the \old{Relative Time}\new{RT} volume 
    can be computed by the predictive painting from the estimated slopes, as 
    discussed in the previous section. 
    \old{Then with}\new{With} the help of the generated RT volume, the RT-seislet transform is 
    \new{then} applied to the input data to get coefficients in the RT-seislet transform 
    domain. 
    With the transformed data, signal processing tasks, including data 
    reconstruction and noise attenuation, can be easily applied.

    \inputdir{.}
    \plot{workflow}{width=0.8\columnwidth}{Workflow of the proposed RT-seislet
    transform.}

%-----------------------------------------------------------------
\section{Synthetic Data Example}
    We use the same data example from Figure~\ref{fig:sigmoid} to demonstrate 
    the performance of the proposed method in ef{}ficient data reconstruction. 
    Figure~\ref{fig:rt} shows the \old{Relative Time}\new{RT} calculated from 
    the image using the predictive painting with multiple reference traces. 
    We applied both the original PWD-seislet transform and the RT-seislet 
    transform to this image. 
    Then, we threshold coefficients and keep only the largest ones from both the 
    transforms to reconstruct the same original data. 
    By using the same percent (1\%) of large coefficients, the reconstruction 
    from the new formulation (Figure \ref{fig:rtseisrec1}) recovers more 
    accurate amplitude values and better preserves structural discontinuities 
    near faults and unconformities than the one from PWD-seislet transform 
    (Figure \ref{fig:pwdseisrec1}).

    \inputdir{sigmoid}
    \plot{rt}{width=0.8\columnwidth}{\old{Relative time}\new{RT} volume 
        estimated by the predictive painting using multiple reference traces.}
    \multiplot{2}{pwdseisrec1,rtseisrec1}{width=0.6\columnwidth}{Synthetic 
        seismic image reconstruction by thresholding coefficients and keeping 
        only 1\% of significant coefficients using (a) inverse PWD-seislet 
        transform and (b) inverse RT-seislet transform.}
    \plot{coef}{width=0.8\columnwidth}{Normalized coefficients from the 
        PWD-seislet transform (red \new{dashed} line) and the RT-seislet 
        transform (blue \new{solid} line) sorted from large to small on a 
        decibel scale.}

    The comparison between coefficients from the PWD-seislet transform and 
    the RT-seislet transform is shown in Figure \ref{fig:coef}. 
    The minor difference of their decay rate indicates that the PWD-seislet 
    transform and the proposed transform obtain the same sparsity for the input 
    data.
    Figure \ref{fig:tim} shows the plot of CPU time against the number of 
    traces. The proposed formulation of the seislet transform shows promising 
    computation efficiency since the prediction of a trace using distant 
    traces is accomplished directly instead of recursively as in the original 
    PWD-seislet transform.

    \inputdir{time}
    \plot{tim}{width=0.80\columnwidth}{CPU time of the PWD-seislet transform 
        (red \new{dashed} line) and the RT-seislet transform (blue \new{solid}
        line) \old{vs.}\new{versus} the number of traces.}

    For the purpose of testing the sensitivity of the RT-seislet transform to 
    noise, we add random noise to the synthetic model and utilize 
    signal-to-noise ratio (SNR) as measurement: 
    \begin{equation}
        \label{snr}
        SNR=10\log_{10}\frac{\|\mathbf{s}_{true}\|^2_2}
        {\|\mathbf{s}_{true}-\mathbf{s}\|^2_2}\,,
    \end{equation}
    where $\mathbf{s}_{true}$ is the original data, and $\mathbf{s}$ denotes
    the reconstructed data (e.g., Figure~\ref{fig:rtseisrec1}) or noisy data.
    We plot SNRs of reconstructed data using the inverse PWD-seislet transform 
    and the inverse RT-seislet transform with respect to the noise variance in 
    Figure~\ref{fig:snr}. 
    Apparently, as the noise level increases, the performance of the RT-seislet 
    gets worse, which shows that the proposed method is sensitive to noise. 
    But, compared with the original PWD-seislet transform, the new method has 
    superior performance \old{with less sensitivity}\new{that is less sensitive} to noise.
    Figure~\ref{fig:sigmoid-noise0,rtseisrec-noise0} shows one example of the
    noisy data and the reconstructed data. The SNR of 
    Figure~\ref{fig:rtseisrec-noise0} is 17.19 dB.

    \inputdir{noise}
    \plot{snr}{width=0.80\columnwidth}{SNR diagram of the PWD-seislet transform 
        (red \new{dashed} line) and the RT-seislet transform (blue \new{solid}
        line) with respect to the noise level (variance value).}
    \multiplot{2}{sigmoid-noise0,rtseisrec-noise0}{width=0.6\columnwidth}{(a)
        Noisy data. (b) Reconstructed data using 10\% coefficients (SNR=17.19).}

    Figure~\ref{fig:para,para-rtseisrec1,para-diff} presents another synthetic example. 
    Figure~\ref{fig:para} and \ref{fig:para-rtseisrec1} shows the synthetic shot gather with three hyperbolic 
    events and Figure~\ref{fig:para-rtseisrec1} is the data reconstruction by the 
    RT-seislet transform using only 1\% of the most significant coefficients, respectively. 
    We use \old{only the middle trace}\new{the first, the middle and the last 
    traces} as the reference trace\new{s} to compute the 
    \old{Relative Time}\new{RT} volume.
    \new{The residual between the synthetic shot gather and the reconstruction
    result is shown in Figure~\ref{fig:para-diff}.}
    For this simple synthetic shot gather, the proposed method achieves an 
    excellent data reconstruction result.

    \inputdir{synth}
    \multiplot{3}{para,para-rtseisrec1,para-diff}{width=0.45\columnwidth}{
        (a) Synthetic shot gather. (b) Data reconstruction by the inverse 
        RT-seislet transform using only 1\% of the most significant 
        coefficients. (c) Difference between (a) and (b).}

\section{2-D Field Data Example}
    To further test the proposed method, we use a common-midpoint gather from the 
    North Sea dataset (Figure~\ref{fig:gath}). Figure~\ref{fig:seis} shows the 
    data in the seislet transform domain. We observe that significant 
    coefficients are apparent in only a narrow area on the left, which 
    indicates the excellent compression ability of the proposed RT formulation 
    of the seislet transform. The field data reconstruction using only 1\% of 
    the most significant RT-seislet coefficients is shown in 
    Figure~\ref{fig:seisrec1}. The reconstructed image \old{contains}\new{retains} the most 
    important seismic events in the original seismic data.

    If we mute the RT-seislet coefficients at fine scales while keeping the 
    significant coefficients at coarse scales (Figure~\ref{fig:seisdeno}), the 
    inverse RT-seislet transform enables effective denoising, removing 
    incoherent noise from the gather. 
    The result of denoising are shown in Figure~\ref{fig:seis16}. 
    Figure~\ref{fig:diff} shows the removed noise section using the proposed implementation of the seislet transform.

    The RT-seislet is also able to perform interpolation. 
    We add two more scales with small random noise to Figure~\ref{fig:seis} and 
    interpolate the \old{Relative Time}\new{RT} volume by a factor of four. 
    Then the interpolated shot gather is obtained and the number of traces is 
    increased by four. 
    The extended data in the seislet transform domain is shown in Figure~\ref{fig:seisrand2}. 
    Figure~\ref{fig:gath4} is the interpolated shot gather. 
    Figure \ref{fig:fft,fft2} shows the F-K spectrum of the zero-padded 
    field data and the interpolated gather. 
    The zero-padded shot gather is generated by manually padding zero traces 
    between two traces.
    In this example, the seislet transform domain is extended with random noise 
    because there is unpredictable noise in the real case.

    \inputdir{gath}
    \plot{gath}{width=0.80\columnwidth}{Common-midpoint gather.}
    \multiplot{2}{seis,seisrec1}{width=0.60\columnwidth}{(a) The field data in 
        the RT-seislet transform domain and (b) \old{the field data }reconstruction using only 1\% of significant coefficients by the inverse 
        RT-seislet transform.}
    \multiplot{4}{gath,seisdeno,seis16,diff}{width=0.45\columnwidth}{\new{(a) Original gather. }(b) \old{Denoising}\new{Denoised} 
        result by muting RT-seislet coefficients at fine scales (a) and (c) 
        the removed noise section using the RT-seislet transform.}
    \multiplot{2}{seisrand2,gath4}{width=0.6\columnwidth}{(a) Extended seislet 
        transform domain. (b) Interpolated shot gather by the inverse RT-seislet 
        transform.}
    \multiplot{2}{fft,fft2}{width=0.6\columnwidth}{F-K \old{spectrum}\new{spectra} of (a) 
        zero-padded \old{field data} and (b) the interpolated field data.}

\section{3-D Field Data Example}
    We also test the proposed method on a 3D dataset. 
    Figure~\ref{fig:cuber} is the Teapot Dome seismic dataset. 
    After inline and crossline slopes (Figure~\ref{fig:dip1,dip2}) are 
    estimated, the \old{Relative Time}\new{RT} volume (Figure~\ref{fig:pick}) is 
    easily obtained from the predictive painting with multiple reference traces. 
    With the RT volume, the 3D seislet transform can be constructed by 
    cascading the 2D seislet transform along inline and crossline directions. 
    To be specific, we first apply the 2D RT-seislet transform along inline 
    direction, and then another 2D transform along crossline is applied on the 
    coefficients from the previous step to obtain the coefficients in the 3D 
    transform domain.
    The coefficients of the 2D seislet transform along inline direction and the 
    3D seislet transform are shown in Figure \ref{fig:rtseis-inline} and 
    \ref{fig:rtseis}.
    In the seislet domain, most of the coefficients concentrate in the coarse 
    scales, which shows the excellent compression ability of the RT-seislet 
    transform. 
    Figure~\ref{fig:teapot-rtseisrec5} is the data reconstruction result using 
    only 5\% of the most significant coefficients. 
    The difference between Figure \ref{fig:cuber} and the reconstructed data, 
    contains mainly noise, is shown in Figure \ref{fig:teapot-diff}.
    The proposed method has \old{excellent}\new{good} performance for reconstruction and 
    preserving the structures of the dataset.

    \inputdir{teapot}
    \multiplot{2}{cuber,pick}{width=0.6\columnwidth}{(a) Teapot dome dataset. 
        (b) \old{Relative time}\new{RT} volume estimated by the predictive 
        painting.}
    \multiplot{2}{dip1,dip2}{width=0.6\columnwidth}{Inline (a) and Crossline (b)
        Dip.}
    \multiplot{4}{rtseis-inline,rtseis,teapot-rtseisrec5,teapot-diff}
    {width=0.4\columnwidth}{(a) The coefficients in 2D seislet domain along 
        inline only. (b) The coefficients in 3D seislet domain. 
        (c) Data reconstruction using only 5\% of significant coefficients by 
        the inverse RT-seislet transform.
        (d) Difference between (c) and the original seismic volume 
        (Figure \ref{fig:cuber}).}

%-----------------------------------------------------------------
\section{Discussion}
    The proposed formulation of the seislet transform attempts to construct 
    multiscale prediction operators, to enhance the performance of the 
    seislet transform. 
    The prediction operators in the original seislet transform are defined by 
    PWD, so the prediction of distant traces needs recursive computation, which 
    is very time-consuming when it comes to large size dataset. 
    In the proposed RT-seislet transform, an RT volume is used to define a 
    multiscale prediction operator. 
    As shown in the previous sections, the relationship between any two traces 
    can be obtained from the \old{Relative Time}\new{RT} volume, which helps 
    avoid the recursive computation for the prediction of distant traces. 
    As a result, the RT-seislet transform is much more efficient than the 
    original seislet transform.

    Besides, in the implementation of the proposed method, the choice of 
    reference traces plays an important role in the generation of the RT volume. 
    More reference traces usually mean a better \old{Relative Time}\new{RT} 
    volume. 
    The location of reference traces also needs to be considered carefully. 
    In practice, evenly distributed reference traces are recommended.
    Besides, complex geologic structures need to be treated differently.
    For example, it is better to put reference traces at each side of faults. 
    Another important issue is about the smoothing radius when estimating local 
    slopes by PWD. 
    For noise-free data, small smoothing radius is enough. 
    The smoothing radius is three for the estimation of Figure~\ref{fig:dip-pad}. 
    And with the increasing of the noise level, larger smoothing radius is 
    needed.

    The proposed RT-seislet transform highly depends on the RT volume. 
    In our method, local slopes by PWD, which are easily 
    affected by noise, are utilized to generate the RT volume, so the new 
    formulation is sensitive to the noise. 
    However, this problem certainly can be solved by improving the PWD or using 
    other methods, which are not sensitive to noise, to generate the RT volume.

%-----------------------------------------------------------------
\section{Conclusions}
    We have proposed a new formulation of the seislet transform, a novel 
    approach to sparse seismic data. 
    In our approach, we use an RT volume to construct multiscale prediction and 
    update operators for the seislet transform. 
    This proposed approach is computationally more efficient and better 
    preserves geologically meaningful discontinuities than the original 
    formulation of the seislet transform in reconstructing the \old{original 
    seismic dataset}\new{raw seismic data}. 
    It also has excellent performance in obtaining the sparse representation of
    seismic data.
    Synthetic \old{data }and field data examples show that the proposed new 
    formulation is efficient and effective in data reconstruction and noise 
    attenuation.

%-----------------------------------------------------------------
\section{Acknowledgements}
    The research is partially supported by Texas Consortium for Computation (TCCS), the “Thousand Youth Talents Plan”, and the Starting Funds from Zhejiang University. The computational examples reported in this paper are reproducible using the \emph{Madagascar} open-source software package \cite[]{Fomel2013}.

\section{DATA AND MATERIALS AVAILABILITY}
Data associated with this research are available and can be obtained by contacting the corresponding author.

\bibliographystyle{seg}
\bibliography{rtseis}